---
title: "Wildfire Relief Data Visualization"
output:
  distill::distill_article:
    toc: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      layout = "l-body-outset",
                      fig.height = 6)
```

```{r load_packages}
library(tidyverse)
library(devtools)
library(leaflet)
library(sf)
library(lubridate)
library(tfff)

load_all()
```

```{r load_data}
wildfires_2020_at_largest_date <- wildfires_2020 %>% 
  st_drop_geometry() %>% 
  mutate(date = as_date(polygon_date_time)) %>% 
  group_by(date) %>% 
  summarize(total_acres = sum(gis_acres, na.rm = TRUE)) %>% 
  drop_na(date) %>% 
  slice_max(total_acres, n = 1) %>% 
  pull(date)

wildfires_2020_at_largest_single_day <- wildfires_2020 %>% 
  mutate(date = as_date(polygon_date_time)) %>% 
  filter(date == wildfires_2020_at_largest_date)

wildfires_2020_at_largest <- wildfires_2020 %>% 
  group_by(incident_name) %>% 
  slice_max(gis_acres, n = 1, with_ties = FALSE) %>% 
  ungroup() 
```


```{r}
map_basemap <- function() {
  leaflet() %>%
    addProviderTiles(providers$CartoDB.Positron) %>% 
  setView(lng = -120.28940551789255, lat = 43.829711619763955, zoom = 6)
}
```


# Wildfires

## Wildfires 2020

This map shows the wildfires at their largest point in 2020. 

```{r}
map_wildfires <- function(leaflet_layers) {
  leaflet_layers %>%
    addPolygons(data = wildfires_2020_at_largest,
                group = "Wildfires",
                color = tfff_colors("Orange"),
                smoothFactor = 0.2,
                fillOpacity = 0.9,
                stroke = FALSE) 
}


map_basemap() %>% 
  map_wildfires()

```

## Wildfires 2020 Compared to Previous Years

I'll add this later on.

# COVID

Darker blue = more cases per capita

## COVID Cases: September 1

```{r}
map_covid <- function(leaflet_layers, 
                      observation_date,
                      group_name = NULL) {
  
  
  pal <- colorNumeric(
    palette = tinter::tinter(tfff_colors("Blue"),
                             steps = 5,
                             direction = "tints"),
    domain = covid_data$cases_seven_day_avg)
  
  leaflet_layers %>% 
    addPolygons(data = filter(covid_data, date == observation_date),
                group = group_name,
                color = ~pal(cases_seven_day_avg),
                smoothFactor = 0.2,
                fillOpacity = 0.5,
                stroke = 0.1) 
}
```

```{r}
map_basemap() %>% 
  map_covid("2020-09-01")  
```


## COVID Cases: October 1

```{r}
map_basemap() %>% 
  map_covid("2020-10-01")  
```

## COVID Cases: November 1

```{r}
map_basemap() %>% 
  map_covid("2020-11-01")  
```


# Communities

The yellow polygons show the borders of communities in Oregon and Siskiyou counties. 

```{r}
map_communities <- function(leaflet_layers) {
  leaflet_layers %>% 
    addPolygons(data = oregon_siskiyou_communities_geospatial,
                group = "Communities",
                color = tfff_colors("Yellow"),
                fillOpacity = 0.9,
                stroke = FALSE,
                popup = ~community)
}

map_basemap() %>% 
  map_communities()
```


## Percent of Mobile Housing

This shows basically the same thing (i.e. yellow = community boundaries), but darker yellow indicates a higher percentage of units that are mobile housing. I broke the communities into thirds so the darkest yellow is the top third of communities, middle third is a medium yellow, bottom third (i.e. lowest percentage of mobile homes) is the lightest yellow. Click on any community to see the percentage of mobile homes.

```{r}
tfff_yellows <- tinter::tinter(tfff_colors("Yellow"),
                             steps = 3,
                             direction = "tints")




map_mobile_housing <- function(leaflet_layers) {
  
  mobile_housing <- mobile_housing %>% 
    mutate(quartile = ntile(pct_mobile_homes, 3)) %>% 
  mutate(fill_color = case_when(
    quartile == 1 ~ tfff_yellows[1],
    quartile == 2 ~ tfff_yellows[2],
    quartile == 3 ~ tfff_yellows[3]
  )) %>% 
    mutate(label = str_glue("<b>{community}</b><br>{(percent(pct_mobile_homes, 1))} of units are mobile homes"))
  
  
  leaflet_layers %>% 
    addPolygons(data = mobile_housing,
                group = "Mobile Housing",
                color = ~fill_color,
                fillOpacity = 0.9,
                stroke = FALSE,
                popup = ~label)
}

map_basemap() %>% 
  map_mobile_housing()
```


## Driving Time to Nearest Hospital

I calculated the driving time from the center of each community to the closest hospital. Then, I plotted each community on the map using the same yellow color scheme from mobile housing (i.e. break communities into thirds, darker yellow = farther from hospital). These times are approximate as we can't calculate this for every address in every community, but it gives a good overall sense. 

```{r}


map_closest_hospital <- function(leaflet_layers) {
  
 
 closest_hospitals_geospatial <- closest_hospitals %>% 
  separate(community, 
           into = c("community", "state"),
           sep = ", ") %>% 
  left_join(oregon_siskiyou_communities_geospatial) %>% 
  mutate(quartile = ntile(minutes_to_hospital, 3)) %>% 
  mutate(fill_color = case_when(
    quartile == 1 ~ tfff_yellows[1],
    quartile == 2 ~ tfff_yellows[2],
    quartile == 3 ~ tfff_yellows[3]
  )) %>% 
   mutate(label = str_glue("<b>{community}</b><br>{number(minutes_to_hospital, 1)} minutes to closest hospital")) %>% 
  st_as_sf()

  
  leaflet_layers %>% 
    addPolygons(data = closest_hospitals_geospatial,
                group = "Driving Time to Nearest Hospital",
                color = ~fill_color,
                fillOpacity = 0.9,
                stroke = FALSE,
                popup = ~label)
}

map_basemap() %>% 
  map_closest_hospital()
```


# Rebuilding

## Hospitals

This shows hospitals in Oregon and Siskiyou. The darker green shows rural hospitals, as taken from the [Oregon Office of Rural Health](https://www.ohsu.edu/media/966) (while Siskiyou hospitals are likely rural as well, I don't have data from California to classify them).

```{r}
map_hospitals <- function(leaflet_layers) {
  
  hospitals <- hospitals %>% 
    mutate(color_for_marker = case_when(
      rural_hospital == "Rural" ~ "green",
      TRUE ~ "darkgreen"
    )) %>% 
    mutate(icon_color = case_when(
      rural_hospital == "Rural" ~ "black",
      TRUE ~ "#ffffff"
    ))
  
  hospital_icon <- awesomeIcons(
    icon = 'h-square',
    iconColor = 'black',
    library = 'fa',
    markerColor = hospitals$color_for_marker
  )
  
  leaflet_layers %>% 
    addAwesomeMarkers(data = hospitals,
                      group = "Hospitals",
                      icon = hospital_icon,
                      popup = ~name) 
}

map_basemap() %>% 
  map_hospitals()
```



## Clinics

At this point, I only have data for clinics from California. If you know of any source for Oregon data, please let me know. I've done some searching and haven't turned up anything.

```{r}
# map_basemap() %>% 
#   addCircleMarkers(data = clinics_california,
#                    popup = ~fac_name) 
```


# Combined Map

This map shows all of the above maps in a single map, with the ability to filter out certain layers. The community-level layers (communities, mobile housing, and distance to nearest hospital) overlap so if you want to see those, you'll want to just select one. 



```{r}
map_basemap() %>% 
  map_wildfires() %>% 
  map_covid("2020-11-01",
            group_name = "COVID Cases (November 1)") %>%
  map_communities() %>% 
  map_hospitals() %>% 
  map_mobile_housing() %>% 
  map_closest_hospital() %>% 
  # hideGroup("Mobile Housing") %>% 
  addLayersControl(
    overlayGroups = c("Wildfires", 
                      "Hospitals", 
                      "Communities", 
                      "Mobile Housing",
                      "Driving Time to Nearest Hospital",
                      "COVID Cases (November 1)"),
    options = layersControlOptions(collapsed = FALSE,
                                   autoZIndex = FALSE))
```


